Spin canonical divisors of log stacky curves

### 2015-06-21 stacky_write_up/spin_cring.tex

\begin{lem}
\label{lem:deg1_sat_ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as defined by O'Dorney\todo{reference this-Section 1 of Evan's paper}) satisfying $\deg(\lfloor{D}\rfloor)\ge 0$, so $R_D$ has a generator $u$ in degree 1; let $u$.  Suppose $D$ is generated by $u,x_1,...,x_m$ in degree at most $\delta$ with relations generated in degree at most $\tau$.  Suppose $D'=D+\frac{\alpha}{\beta}P$ for some $P\not\in \Supp(D), \alpha,\beta\in \mathbf{N}$ such that 
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X,\lfloor{dD+d\frac{\alpha}{\beta}P\rfloor})=h^0(X,\lfloor dD
	\rfloor)+\lfloor d\frac{\alpha}{\beta}\rfloor \forall d\in \mathbf{N}:
	d\ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  degree $d_i$ and has a pole at $P$ of degree $c_i$
\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then 
\begin{align*}
	\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle y_i x_j: 1 \le i \le n, 1 \le j \le m \rangle \\
										 &+ \langle y_i y_j: 1 \le i \le j \le n-1 \rangle.
\end{align*}
\todo{See if we can additionally get the other information discussed in VZB theorem 8.3.1}
\end{itemize}
\end{lem}

\begin{proof}
Observe that $u_D,x_1,...,x_m$ each of which has no pole at $P$.  These can be chosen generically to not have a zero at $P$ \todo{check that we can actually do this, or eliminate if it turns out to be unnecessary}.

Begin by focusing on the $d^{\text{th}}$ component of $R_{D'}$: $(R_
{D'})_d = H^0(X, \lfloor D + \frac{\alpha}{\beta} P \rfloor)$.  By our
hypothesis (Equation ~\ref{eqn:deg1_sat_ind_dim}), for any $d\in\mathbf{N}$ such that $\lfloor d\frac{
\alpha}{\beta}\rfloor>0$,

@@

Let $0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{
\beta}$ be the non-negative best lower approximation of $\frac{
\alpha}{\beta}$.  Note that elements of $H^0(X,dD')$ have pole of
orders ranging from 0 to $\lfloor d_i \frac{\alpha}{\beta}\rfloor =
c_i $ at $P$.  For any $j < i$, $\frac{c_i}{d_i} > \frac{c_j}{d_j}$
by construction, so any product of $\prod z_i$ of elements of
degrees less than $d_i$ must have $\frac{-\ord(z_i)}{\deg(z_i)} <
\frac{c_i}{d _i}$, so we have the pole order of this product is $
\sum -\ord(z_i) = d_i(\frac{\sum -\ord( z_i)}{\sum \deg( z_i)}) < d_
i\frac{c_i}{d_i} = c_i $. Thus, the elements of $R_{d_i}$

In fact, a generic element in the respective degree will suffice.
\todo{add proof of this; is the generic statement correct?}
Following the argument given by O'Dorney \todo{cite, give specific
theorem number} correctly gives that these along with $u$ (in
degree 1 with 0 pole at $P$) generate a $\lfloor d\frac{ \alpha}{
\beta}\rfloor+1$ dimension $k$-vector space inside of $(R_{ D'})_d$
with a $\lfloor d\frac{\alpha}{\beta}\rfloor$ dimensional subspace
consisting of elements only with nonzero pole at $P$. Therefore, by
pole considerations $y_1,...,y_n$ generate $(R_{D'})_d $ over $(R_{D
})_d$, \todo{the word generate here is a little imprecise} so $u,x_1
,...,x_m,y_1,...,y_n$ generate $R_{D'}$.

@@

>>>>

\begin{lem}
\label{lem:deg1_sat_ind}
Let $X$ be a genus $g$ curve with a $\mathbf{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper})
satisfying $\deg(\lfloor{D}\rfloor)\ge 0$, so $R_D$ has a
generator $u$ in degree 1; let $u$.  Suppose $D$ is generated by $
u, x_1, \ldots, x_m$ in degree at most $\delta$ with relations generated 
in degree at most $\tau $.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $P \not \in \Supp(D), \alpha,\beta \in \mathbf{N}$ such
that

\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \forall d \in \mathbf{
	N} : d \ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$
\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then

\begin{align*}
	\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle y_i x_j: 1 \le i \le n, 1 \le j \le m \rangle \\
										 &+ \langle y_i y_j: 1 \le i \le j \le n-1 \rangle.
\end{align*}

\todo{See if we can additionally get the other information discussed in VZB theorem 8.3.1}
\end{itemize}
\end{lem}

\begin{proof}
Observe that $u_D,x_1,...,x_m$ each of which has no pole at $P$.
These can be chosen generically to not have a zero at $P$ \todo{
check that we can actually do this, or eliminate if it turns out to
be unnecessary}.

Begin by focusing on the $d^{\text{th}}$ component of $R_{D'}$: $(R_
{D'})_d = H^0(X, \lfloor D + \frac{\alpha}{\beta} P \rfloor)$. By
our hypothesis (Equation ~\ref{eqn:deg1_sat_ind_dim}), for any $d
\in \mathbf{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

Let $0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{
\beta}$ be the non-negative best lower approximation of $\frac{
\alpha}{\beta}$.  Note that elements of $H^0(X,dD')$ have pole of
orders ranging from 0 to $\lfloor d_i \frac{\alpha}{\beta}\rfloor =
c_i $ at $P$. A set of these one for each pole degree 1 to $\lfloor
d_i \frac{\alpha}{\beta} \rfloor$ are linearly independent from
each other and $(R_D)_d$, so by degree considerations, they span
$(R_{D'})_d$ over $(R_D)_d$ as $k$-vector spaces.

For any $j < i$, $\frac{c_i}{d_i} > \frac{c_j}{d_j}$
by construction, so any product of $\prod z_i$ of elements of
degrees less than $d_i$ must have $\frac{-\ord(z_i)}{\deg(z_i)} <
\frac{c_i}{d _i}$, so we have the pole order of this product is $
\sum -\ord(z_i) = d_i(\frac{\sum -\ord( z_i)}{\sum \deg( z_i)}) < d_
i\frac{c_i}{d_i} = c_i $. Thus, the elements of $R_{d_i}$

For any $j<i$, $\frac{c_i}{d_i}>\frac{c_j}{d_j}$ by construction,
so any product of $\prod z_i$ of elements of degrees less than $d_i$
 must have $\frac{-\ord(z_i)}{\deg(z_i)}<\frac{c_i}{d_i} $, so we
have the pole order of this product is $\sum -\ord(z_i)=d_i(\frac{
\sum -\ord( z_i)} {\sum \deg( z_i)})<d_i\frac{c_i}{d_i}=c_i$.
Thus, elements of $(R_{D'})d_i$ with pole order $c_i$ at $P$ are
not generated by lower degrees; generically choose some $y_i \in(R_{
D'})_{d_i}$ to have maximal pole order (note that if $d_i=1$ then $-
\ord_P(y_i)= \max(c_j:d_j=1)$ and if more than one $d_i$ is 1, we
must also choose them generically so they generate $(R_{D'})_1$).

Suppose $d\le \beta$ and $\frac{\bar{c}}{\bar{d}}\le \frac{\alpha}{
\beta}$ with $c>0$ is not a best lower approximation.  Then choose
best lower approximation $\frac{c_i}{d_i}$ such that $c_i\le c$ is
maximal; since $c<c_{i+1}$, $\frac{c_i}{d_i}\ge \frac{c}{d}$.  From
this we can deduce that $c_i+(d-d_i)\lfloor \frac{\alpha}{\beta}
\rfloor\ge c$.  Therefore, we can produce an element in $(R_{D'})_d$
 with pole degree $c$ at $P$ by either $y_i u^b$ or $y_j y_i u^b$
where $d_j=1$ and if $d_i=1$ as well then $c_i=1$. Thus, for a
general $d$ (not necessarily less than $\beta$), if $\frac{c}{d} \le
\frac{\alpha}{\beta}$ is not a best lower approximation than we can
construct an element in degree $d$ with pole degree $c$ by either $y
_n^a y_i u^b$ or $y_n^a y_j y_iu^b$ with the latter satisfying the
same conditions.  We notice that elements of this form are all
linearly independent in $R_{D'}$.  Thus, they form a $k$-basis for
$R_{D'}$ over $R_D$.  Thus $u,x_1,...,x_m,y_1,...,y_n$ generate
$R_{D'}$.

\todo{finish proof of relations, which follows easily using VZB's
method; peter will do this tomorrow (sunday)}

@@

\todo{lots of corollaries; peter can do this tomorrow}
